---
title: "PracticalML"
author: "Raymond Wu"
date: "9/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This is for practical machine learning course project. 

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: <http://groupware.les.inf.puc-rio.br/har> (see the section on the Weight Lifting Exercise Dataset)

#### Data
The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

Data from <http://groupware.les.inf.puc-rio.br/har>. It's very generous allowing their data to be used for this project. 
Source from : 
Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

#### Loading packages

```{r message=FALSE}
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
library(rattle)
```

#### Setup working directory
Download training and testing data to working directory
```{r message=FALSE}
setwd("C:/CourseProject")
trainFileName <- "pml-training.csv"
testFileName <- "pml-testing.csv"
```

#### Data cleansing
Observed training and testing data file contains "NA" and blank in some data item. Some data items contains "#DIV/0!" in training data file. These are not desired items
The first 7 columns are non-predictors 
```{r message=FALSE}
trainSource <- read.csv(trainFileName, na.strings=c("NA","","#DIV/0!"))
testSource <- read.csv(testFileName, na.strings=c("NA","","#DIV/0!"))
trainData <- trainSource[, colSums(is.na(trainSource)) == 0]
testData <- testSource[, colSums(is.na(testSource)) == 0]
trainData <- trainData[, -c(1:7)]
testData <- testData[, -c(1:7)]
dim(trainData)
dim(testData)
```

#### Try to remove zero covariates in training data
```{r message=FALSE}
NZV <- nearZeroVar(trainData[,-ncol(trainData)],saveMetrics = TRUE)
```
Verify the result
```{r message=FALSE}
NZV
```
These are features after cleansing training data
```{r message=FALSE}
rownames(NZV)
```

## Partitioning training data to train set/ test set
```{r message=FALSE}
set.seed(1235)
inTrain <- createDataPartition(trainData$classe, p = 0.75, list = FALSE)
inTraining <- trainData[inTrain, ]
inTesting <- trainData[-inTrain, ]
dim(inTraining)
dim(inTesting)
```
## Random forest model
#### choosing 4-fold cross validation 
```{r message=FALSE}
trCtrRF <- trainControl(method="cv", number = 4)
modelRF <- train(classe ~., data = inTraining, method = "rf", metric = "Accuracy", trControl=trCtrRF )

modelRF$finalModel

predictRF <- predict(modelRF , newdata = inTesting)

cmRF <- confusionMatrix(predictRF, as.factor(inTesting$classe))
cmRF

varImp(modelRF)
```
- Accuracy rate is very high. it is 0.9929, which means out-of-sample-error is 0.0071 or 0.71%  

## Decision tree model
```{r message=FALSE}
set.seed(1235) 
modelDT <- rpart(classe ~., data =inTraining, method="class" )
fancyRpartPlot(modelDT)
predictDT <- predict(modelDT, inTesting, type= "class")
cmDT <- confusionMatrix(predictDT, as.factor(inTesting$classe))
cmDT 
```
- Accuracy rate is 0.7382, which means out-of-sample-error is 0.2618 or 26.18% 

## Choose model to predict test data

Random forest is the one to predict test data since it shows better accuracy

```{r message=FALSE}
preditQuiz <- predict(modelRF, newdata = testData)
preditQuiz
```
